---
title: "Format Michalis' output for explorer"
author: "Joe Hasell"
date: "21/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RCurl)
library(knitr)

library(googlesheets4)

```

Load functions
```{r}
source("scripts/explorer_controls_builder.R")

```

##Prep the data

Read in Michalis' povcal data
```{r}
# df <- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/MichalisMoatsos/PovcalNetMainDataCountriesOnly.csv", row.names = 1, header = T, sep = ";", dec = ",")

df <- read.csv("PovcalNetMainDataCountriesOnly.csv", row.names = 1, header = T, sep = ";", dec = ",")


```

Map 3-letter codes to country names
```{r}
name_mapping <- read.csv("entity_names_three_letter_country_standardized.csv") %>%
  rename(Entity = Country)

df<- left_join(df,name_mapping)

df<- df %>%
  select(-Entity) %>%
  rename(Entity = Our.World.In.Data.Name) %>%
  select(Entity, everything())

```

Rename vars (We should do rearrange Michalis' script to structure the variables like this and then delete this section)

```{r}


orig_varnames<- c(

"Entity",
"Year",
"X.1.per.day...share.of.population.below.poverty.line",
"X.1.per.day...poverty.gap.index",
"X.1.per.day...total.number.of.people.below.poverty.line",
"X.1.per.day...absolute.poverty.gap",
"X.1.per.day...income.gap.ratio",
"X.1.per.day...watts.index",
"X.1.9.per.day...share.of.population.below.poverty.line",
"X.1.9.per.day...poverty.gap.index",
"X.1.9.per.day...total.number.of.people.below.poverty.line",
"X.1.9.per.day...absolute.poverty.gap",
"X.1.9.per.day...income.gap.ratio",
"X.1.9.per.day...watts.index",
"X.3.2.per.day...share.of.population.below.poverty.line",
"X.3.2.per.day...poverty.gap.index",
"X.3.2.per.day...total.number.of.people.below.poverty.line",
"X.3.2.per.day...absolute.poverty.gap",
"X.3.2.per.day...income.gap.ratio",
"X.3.2.per.day...watts.index",
"X.5.5.per.day...share.of.population.below.poverty.line",
"X.5.5.per.day...poverty.gap.index",
"X.5.5.per.day...total.number.of.people.below.poverty.line",
"X.5.5.per.day...absolute.poverty.gap",
"X.5.5.per.day...income.gap.ratio",
"X.5.5.per.day...watts.index",
"X.10.per.day...share.of.population.below.poverty.line",
"X.10.per.day...poverty.gap.index",
"X.10.per.day...total.number.of.people.below.poverty.line",
"X.10.per.day...absolute.poverty.gap",
"X.10.per.day...income.gap.ratio",
"X.10.per.day...watts.index",
"X.15.per.day...share.of.population.below.poverty.line",
"X.15.per.day...poverty.gap.index",
"X.15.per.day...total.number.of.people.below.poverty.line",
"X.15.per.day...absolute.poverty.gap",
"X.15.per.day...income.gap.ratio",
"X.15.per.day...watts.index",
"X.20.per.day...share.of.population.below.poverty.line",
"X.20.per.day...poverty.gap.index",
"X.20.per.day...total.number.of.people.below.poverty.line",
"X.20.per.day...absolute.poverty.gap",
"X.20.per.day...income.gap.ratio",
"X.20.per.day...watts.index",
"X.30.per.day...share.of.population.below.poverty.line",
"X.30.per.day...poverty.gap.index",
"X.30.per.day...total.number.of.people.below.poverty.line",
"X.30.per.day...absolute.poverty.gap",
"X.30.per.day...income.gap.ratio",
"X.30.per.day...watts.index",
"X.40.per.day...share.of.population.below.poverty.line",
"X.40.per.day...poverty.gap.index",
"X.40.per.day...total.number.of.people.below.poverty.line",
"X.40.per.day...absolute.poverty.gap",
"X.40.per.day...income.gap.ratio",
"X.40.per.day...watts.index",
"X40..of.median.income...share.of.population.below.poverty.line",
"X40..of.median.income...poverty.gap.index",
"X40..of.median.income...total.number.of.people.below.poverty.line",
"X40..of.median.income...absolute.poverty.gap",
"X40..of.median.income...income.gap.ratio",
"X40..of.median.income...watts.index",
"X50..of.median.income...share.of.population.below.poverty.line",
"X50..of.median.income...poverty.gap.index",
"X50..of.median.income...total.number.of.people.below.poverty.line",
"X50..of.median.income...absolute.poverty.gap",
"X50..of.median.income...income.gap.ratio",
"X50..of.median.income...watts.index",
"X60..of.median.income...share.of.population.below.poverty.line",
"X60..of.median.income...poverty.gap.index",
"X60..of.median.income...total.number.of.people.below.poverty.line",
"X60..of.median.income...absolute.poverty.gap",
"X60..of.median.income...income.gap.ratio",
"X60..of.median.income...watts.index",
"MeanPerMonth",
"MeanPerMonth_estimated",
"MedianPerMonth",
"MedianPerMonth_estimated",
"PPP",
"Population",
"Decile.1...share.of.income.or.consumption",
"Decile.2...share.of.income.or.consumption",
"Decile.3...share.of.income.or.consumption",
"Decile.4...share.of.income.or.consumption",
"Decile.5...share.of.income.or.consumption",
"Decile.6...share.of.income.or.consumption",
"Decile.7...share.of.income.or.consumption",
"Decile.8...share.of.income.or.consumption",
"Decile.9...share.of.income.or.consumption",
"Decile.10...share.of.income.or.consumption",
"Decile.1...threshold.of.income.or.consumption.per.month",
"Decile.2...threshold.of.income.or.consumption.per.month",
"Decile.3...threshold.of.income.or.consumption.per.month",
"Decile.4...threshold.of.income.or.consumption.per.month",
"Decile.5...threshold.of.income.or.consumption.per.month",
"Decile.6...threshold.of.income.or.consumption.per.month",
"Decile.7...threshold.of.income.or.consumption.per.month",
"Decile.8...threshold.of.income.or.consumption.per.month",
"Decile.9...threshold.of.income.or.consumption.per.month",
"Decile.1...average.income.or.consumption.per.month",
"Decile.2...average.income.or.consumption.per.month",
"Decile.3...average.income.or.consumption.per.month",
"Decile.4...average.income.or.consumption.per.month",
"Decile.5...average.income.or.consumption.per.month",
"Decile.6...average.income.or.consumption.per.month",
"Decile.7...average.income.or.consumption.per.month",
"Decile.8...average.income.or.consumption.per.month",
"Decile.9...average.income.or.consumption.per.month",
"Decile.10...average.income.or.consumption.per.month",
"Gini.index",
"Gini_estimated",
"Polarization",
"Polarization_estimated",
"MLD",
"MLD_estimated",
"Palma",
"P90.P10.ratio",
"P90.50.ratio",
"Entropy_0_5",
"Entropy_1_0",
"Entropy_1_5",
"Entropy_2_0",
"Atkinson_0_5",
"Atkinson_1_0",
"Atkinson_1_5",
"Atkinson_2_0",
"Theil_0_0",
"Theil_0_5",
"Theil_1_0",
"Theil_1_5",
"Theil_2_0",
"Var.Coeff",
"datayear",
"isinterpolated",
"usemicrodata",
"coveragetype",
"datatype",
"IsSurveyYear",
"OriginalMedian",
"OriginalMean",
"OriginalDecileShares",
"MonotonicityBreaks",
"RowsWithIncreasingHeadcount",
"DataframeRowsForGpinter",
"LessThan33Rows",
"ISO3DataYearCovType",
"ISO3Year"
)


new_varnames<- c(

"Entity",
"Year",
"headcount_ratio_1_00",
"poverty_gap_index_1_00",
"headcount_1_00",
"total_shortall_annual_1_00",
"income_gap_ratio_1_00",
"watts_index_1_00",
"headcount_ratio_1_90",
"poverty_gap_index_1_90",
"headcount_1_90",
"total_shortall_annual_1_90",
"income_gap_ratio_1_90",
"watts_index_1_90",
"headcount_ratio_3_20",
"poverty_gap_index_3_20",
"headcount_3_20",
"total_shortall_annual_3_20",
"income_gap_ratio_3_20",
"watts_index_3_20",
"headcount_ratio_5_50",
"poverty_gap_index_5_50",
"headcount_5_50",
"total_shortall_annual_5_50",
"income_gap_ratio_5_50",
"watts_index_5_50",
"headcount_ratio_10_00",
"poverty_gap_index_10_00",
"headcount_10_00",
"total_shortall_annual_10_00",
"income_gap_ratio_10_00",
"watts_index_10_00",
"headcount_ratio_15_00",
"poverty_gap_index_15_00",
"headcount_15_00",
"total_shortall_annual_15_00",
"income_gap_ratio_15_00",
"watts_index_15_00",
"headcount_ratio_20_00",
"poverty_gap_index_20_00",
"headcount_20_00",
"total_shortall_annual_20_00",
"income_gap_ratio_20_00",
"watts_index_20_00",
"headcount_ratio_30_00",
"poverty_gap_index_30_00",
"headcount_30_00",
"total_shortall_annual_30_00",
"income_gap_ratio_30_00",
"watts_index_30_00",
"headcount_ratio_40_00",
"poverty_gap_index_40_00",
"headcount_40_00",
"total_shortall_annual_40_00",
"income_gap_ratio_40_00",
"watts_index_40_00",
"headcount_ratio_40_median",
"poverty_gap_index_40_median",
"headcount_40_median",
"total_shortall_annual_40_median",
"income_gap_ratio_40_median",
"watts_index_40_median",
"headcount_ratio_50_median",
"poverty_gap_index_50_median",
"headcount_50_median",
"total_shortall_annual_50_median",
"income_gap_ratio_50_median",
"watts_index_50_median",
"headcount_ratio_60_median",
"poverty_gap_index_60_median",
"headcount_60_median",
"total_shortall_annual_60_median",
"income_gap_ratio_60_median",
"watts_index_60_median",
"MeanPerMonth",
"MeanPerMonth_estimated",
"MedianPerMonth",
"MedianPerMonth_estimated",
"PPP",
"Population",
"share_decile_1",
"share_decile_2",
"share_decile_3",
"share_decile_4",
"share_decile_5",
"share_decile_6",
"share_decile_7",
"share_decile_8",
"share_decile_9",
"share_decile_10",
"threshold_annual_decile_1",
"threshold_annual_decile_2",
"threshold_annual_decile_3",
"threshold_annual_decile_4",
"threshold_annual_decile_5",
"threshold_annual_decile_6",
"threshold_annual_decile_7",
"threshold_annual_decile_8",
"threshold_annual_decile_9",
"average_annual_decile_1",
"average_annual_decile_2",
"average_annual_decile_3",
"average_annual_decile_4",
"average_annual_decile_5",
"average_annual_decile_6",
"average_annual_decile_7",
"average_annual_decile_8",
"average_annual_decile_9",
"average_annual_decile_10",
"Gini",
"Gini_estimated",
"Polarization",
"Polarization_estimated",
"MLD",
"MLD_estimated",
"Palma",
"P90_P10_ratio",
"P90_50_ratio",
"Entropy_0_5",
"Entropy_1_0",
"Entropy_1_5",
"Entropy_2_0",
"Atkinson_0_5",
"Atkinson_1_0",
"Atkinson_1_5",
"Atkinson_2_0",
"Theil_0_0",
"Theil_0_5",
"Theil_1_0",
"Theil_1_5",
"Theil_2_0",
"Var.Coeff",
"datayear",
"isinterpolated",
"usemicrodata",
"coveragetype",
"datatype",
"IsSurveyYear",
"OriginalMedian",
"OriginalMean",
"OriginalDecileShares",
"MonotonicityBreaks",
"RowsWithIncreasingHeadcount",
"DataframeRowsForGpinter",
"LessThan33Rows",
"ISO3DataYearCovType",
"ISO3Year"
)

varname_map<- data.frame(orig_varnames,new_varnames)


```


Swap in new varnames
```{r}
df<- df %>%
  pivot_longer(cols = !c(Entity, Year, coveragetype, datatype, ISO3Year, ISO3DataYearCovType, IsSurveyYear), names_to = "orig_varnames", values_to = "value")

# Merge in new var names

df<- left_join(df, varname_map)

# check there are no unmatched vars
br<- df %>%
  filter(is.na(new_varnames))

# drop original varnames
df<- df %>%
  select(-orig_varnames)

```




Demark Urban/Rural entities.
```{r}
df<- df %>%
  mutate(Entity = if_else(coveragetype == "R",
                          paste0(Entity, " -  Rural"),
                          Entity)) %>%
  mutate(Entity = if_else(coveragetype == "U",
                          paste0(Entity, " -  Urban"),
                          Entity))

```

Change units of some vars
```{r}
# Convert shares into percentages  
df<- df %>%
  mutate(value = if_else(grepl("headcount_ratio|income_gap_ratio|poverty_gap_index",new_varnames),
                         value*100,
                         value))

```

Add calculated vars
```{r}
new_vars<- list()


# Duplicate headcount and headcount ratio vars as a hacky way to have shorter names when shown in a multimetric chart with lots of povlines
new_vars[["headcount_alts"]]<- df %>%
  filter(grepl("headcount_",new_varnames) & !grepl("median",new_varnames)) %>%
  mutate(new_varnames = str_replace(new_varnames, "headcount_", "multi_headcount_"))
         

# Calculate daily decile averages and thresholds from annual values
new_vars[["threshold_annual"]]<- df %>%
  filter(grepl("threshold_annual",new_varnames)) %>%
  mutate(new_varnames = str_replace(new_varnames, "threshold_annual", "threshold_daily"),
         value = value/365)

new_vars[["average_annual"]]<- df %>%
  filter(grepl("average_annual",new_varnames)) %>%
  mutate(new_varnames = str_replace(new_varnames, "average_annual", "average_daily"),
         value = value/365)


# Calculate daily and annual means from monthly values
new_vars[["mean_daily"]]<- df %>%
  filter(new_varnames== "MeanPerMonth") %>%
  mutate(new_varnames = str_replace(new_varnames, "MeanPerMonth", "mean_daily"),
         value = value/(365/12))

new_vars[["mean_annual"]]<- df %>%
  filter(new_varnames== "MeanPerMonth") %>%
  mutate(new_varnames = str_replace(new_varnames, "MeanPerMonth", "mean_annual"),
         value = value*12)

# Calculate daily and annual median from monthly values
new_vars[["median_daily"]]<- df %>%
  filter(new_varnames== "MedianPerMonth") %>%
  mutate(new_varnames = str_replace(new_varnames, "MedianPerMonth", "median_daily"),
         value = value/(365/12))

new_vars[["median_annual"]]<- df %>%
  filter(new_varnames== "MedianPerMonth") %>%
  mutate(new_varnames = str_replace(new_varnames, "MedianPerMonth", "median_annual"),
         value = value*12)

  
# Calculate daily shortfall from annual shortfall
new_vars[["daily_total_shortfall"]]<- df %>%
  filter(grepl("total_shortall_annual",new_varnames)) %>%
  mutate(new_varnames = str_replace(new_varnames, "total_shortall_annual", "total_shortall_daily"),
         value = value/365)
  

# Calculate absolute shortfall per person in poverty (absolute poverty lines only)
new_vars[["avg_shortfall"]]<- df %>%
  filter(grepl("income_gap_ratio",new_varnames) & !grepl("median",new_varnames)) %>%
  mutate(pov_line_text = str_replace(new_varnames, "income_gap_ratio_", "")) %>% #grab poverty line from varname
  mutate(pov_line_math = as.numeric(str_replace(pov_line_text, "_", "."))) %>% #convert povline to numeric
  mutate(new_varnames = str_replace(new_varnames, "income_gap_ratio", "avg_shortfall"),
         value = value/100 * pov_line_math) %>%
  select(-c(pov_line_text, pov_line_math))





# calculate numbers in poverty between pov lines for stacked area
 headcount_stacked_calcs<- df %>%
  filter(grepl("headcount_",new_varnames) & !grepl("ratio",new_varnames)  & !grepl("median",new_varnames)) %>%
  mutate(pov_line_text = str_replace(new_varnames, "headcount_", "")) %>% #grab poverty line from varname
  mutate(pov_line_math = as.numeric(str_replace(pov_line_text, "_", "."))) #convert povline to numeric
  
  # An ordered dataframe of povlines and their variable stubs
   ordered_povlines<- headcount_stacked_calcs %>%
    select(pov_line_text, pov_line_math) %>%
    unique() %>%
    arrange(pov_line_math)
   
   
   selected_povlines<- c(1,1.90, 3.20, 5.50, 10, 30)
   
   ordered_povlines<- ordered_povlines %>%
     filter(pov_line_math %in% selected_povlines)

   
   # create a list for these stacked vars
   stacked_headcounts<- list()

   # Go through each poverty line...
   for(i in 1: length(ordered_povlines$pov_line_math)){
     
     this_povline_stub<- ordered_povlines$pov_line_text[i]
     povline_below_stub<- ordered_povlines$pov_line_text[i-1] #empty if 1=1
  
     # if it's the lowest (first) poverty line...
     if(i==1){
       
       # create a 'stacked_below' variable equal to the headcount
       stacked_headcounts[[paste0("headcount_stacked_below_", this_povline_stub)]]<- df %>%
         filter(new_varnames == paste0("headcount_", this_povline_stub)) %>%
         mutate(new_varnames = str_replace(new_varnames, "headcount_", "headcount_stacked_below_"))
    
     # else if it's the highest (last) poverty line...     
     } else if(i==length(ordered_povlines$pov_line_math)){
       
       # create a 'stacked_above' variable equal to the population minus headcount
       
       df_upper<- df %>%
         filter(new_varnames == "Population") %>%
         rename(upper=value) %>%
         select(-new_varnames)
       
       df_lower<- df %>%
         filter(new_varnames == paste0("headcount_", this_povline_stub)) %>%
         rename(lower=value) %>%
         select(-new_varnames)
       
       df_calc<- merge(df_upper, df_lower) %>%
         mutate(value = upper-lower,
                new_varnames = paste0("headcount_stacked_above_", this_povline_stub)) %>%
         select(-c(upper, lower))
       
       stacked_headcounts[[paste0("headcount_stacked_above_", this_povline_stub)]]<- df_calc
       
       
       # create a 'stacked_below' variable equal to the headcount minus headcount of poverty line below
       
       df_upper<- df %>%
         filter(new_varnames == paste0("headcount_", this_povline_stub)) %>%
         rename(upper=value) %>%
         select(-new_varnames)
       
       df_lower<- df %>%
         filter(new_varnames == paste0("headcount_", povline_below_stub)) %>%
         rename(lower=value) %>%
         select(-new_varnames)
       
       df_calc<- merge(df_upper, df_lower) %>%
         mutate(value = upper-lower,
                new_varnames = paste0("headcount_stacked_below_", this_povline_stub)) %>%
         select(-c(upper, lower))
       
       stacked_headcounts[[paste0("headcount_stacked_below_", this_povline_stub)]]<- df_calc
      
      # else it will be an in between poverty line, in which case... 
     } else{
     
      # create a 'stacked_below' variable equal to the headcount minus headcount of poverty line below
         df_upper<- df %>%
         filter(new_varnames == paste0("headcount_", this_povline_stub)) %>%
         rename(upper=value) %>%
         select(-new_varnames)
       
       df_lower<- df %>%
         filter(new_varnames == paste0("headcount_", povline_below_stub)) %>%
         rename(lower=value) %>%
         select(-new_varnames)
       
       df_calc<- merge(df_upper, df_lower) %>%
         mutate(value = upper-lower,
                new_varnames = paste0("headcount_stacked_below_", this_povline_stub)) %>%
         select(-c(upper, lower))
       
       stacked_headcounts[[paste0("headcount_stacked_below_", this_povline_stub)]]<- df_calc
       
     }
   }

   
  # Collapse list of new stacked_headcounts vars into a single data frame and append to main data
  df_stacked_headcount_vars<- bind_rows(stacked_headcounts)  

  df<- bind_rows(df, df_stacked_headcount_vars)

  

  # calculate shares of population and append to main data
   df_pop<- df %>%
         filter(new_varnames == "Population") %>%
         rename(Population=value) %>%
         select(-new_varnames)
  
  df_stacked_shares_vars<- left_join(df_stacked_headcount_vars, df_pop) %>%
    mutate(value = value/Population * 100) %>%
    mutate(new_varnames = str_replace(new_varnames, "headcount_stacked", "headcount_ratio_stacked")) %>%
    select(-Population)
    
   df<- bind_rows(df, df_stacked_shares_vars)

   
# Collapse list of new vars into a single data frame and append  to main data
df_new_vars_together<- bind_rows(new_vars)

df<- bind_rows(df, df_new_vars_together)


   
```

```{r}

br<- df %>%
  filter(grepl("stacked", new_varnames))
```


Specify which variables to keep for a list of explorers
```{r}

# explorer_var_sets<- list()
# 
# 
# poverty_var_stubs<- c(
#                       "headcount_ratio",
#                       "headcount",
#                       "multi_headcount_ratio",
#                       "multi_headcount",
#                       "headcount_ratio_stacked_below",
#                       "headcount_stacked_below",
#                       "avg_shortfall",
#                       "income_gap_ratio",
#                       "total_shortall_daily",
#                       "poverty_gap_index")
# 
# pov_lines<- c(
#               "1_00",
#               "1_90",
#               "3_20",
#               "5_50",
#               "10_00",
#               "15_00",
#               "20_00",
#               "30_00",
#               "40_00",
#               "40_median",
#               "50_median",
#               "60_median"
# )
# 
# 
# # Make a dataframe with all possible pairings of vars and povlines
# pov_stub_line_combos<- expand.grid(poverty_var_stubs = poverty_var_stubs, 
#                     pov_lines = pov_lines)
#   
# # Drop the avg_shortfall var for relative poverty lines (I wasn't able to calucalate these above -  It is something that should be added to Michalis' script. When that is done, we can remove this exclusion)
# pov_stub_line_combos<- pov_stub_line_combos %>%
#   filter(!(grepl("_median", pov_lines) & poverty_var_stubs == "avg_shortfall"))
# 
# 
# # join the var and povline stubs
# explorer_var_sets[["poverty"]]<- paste0(pov_stub_line_combos$poverty_var_stubs, "_", pov_stub_line_combos$pov_lines)
# 
# 
# # Add the 'stacked_above' vars
# explorer_var_sets[["poverty"]]<- c(explorer_var_sets[["poverty"]],
#                                    "headcount_ratio_stacked_above_30_00",
#                                    "headcount_stacked_above_30_00")
#                                    
#   
# 
#   # And so on for other 
# # explorer_vars[["across_distribution"]]<- c(
# #   
# # )
# 
# 
#   

```


For each explorer, filter the data for the specified vars, flip back to wide format, and then split by survey data type and save as separate csvs
```{r}

#  # Get a vector of names of the different explorers
#   explorers<- names(explorer_var_sets)
# 
# for(explorer in explorers){
#   
#   # Filter for specified vars (NB df is still in long format)
#   df_explorer<- df %>% filter(new_varnames %in% explorer_var_sets[[explorer]])
# 
#   # Flip to long wide format
#   df_explorer<- df_explorer %>%pivot_wider(names_from = "new_varnames", values_from = "value")
# 
#   # Split by survey data type and save separately
#   df_by_survey_type<- list()
#   
# 
#     # Income only, including extrapolated years
#     df_by_survey_type[["inc_extrap"]]<- df_explorer %>%
#       filter(datatype == "income")
# 
#     # Income only, excluding extrapolated years
#     df_by_survey_type[["inc_no_extrap"]]<- df_explorer %>%
#       filter(IsSurveyYear == TRUE,
#              datatype == "income")
# 
#     # Consumption only, including extrapolated years
#     df_by_survey_type[["cons_extrap"]]<- df_explorer %>%
#       filter(datatype == "consumption")
#     
#     # Consumption only, excluding extrapolated years
#     df_by_survey_type[["cons_no_extrap"]]<- df_explorer %>%
#       filter(IsSurveyYear == TRUE,
#              datatype == "consumption")
#     
#     # Both income and consumption, including extrapolated years
#     df_by_survey_type[["inc_or_cons_extrap"]]<- df_explorer
#     
#     # Both income and consumption, excluding extrapolated years
#     df_by_survey_type[["inc_or_cons_no_extrap"]]<- df_explorer %>%
#       filter(IsSurveyYear == TRUE)
#     
#     
#   # Get a vector of names of the different survey types
#   survey_types<- names(df_by_survey_type)
#   
#   # save the separated csvs
#   for(survey_type in survey_types){
#     
#   filename<- paste0(survey_type,".csv")
#   fp<- paste0("data/", explorer, "/", filename)
#   
#   write.csv(df_by_survey_type[[survey_type]], fp, row.names = FALSE)
#   
#   }
#   
# 
#   
# }
# 


```

flip back to wide format, and then split by survey data type and save as separate csvs
```{r}

  # Flip to long wide format
  df<- df %>%pivot_wider(names_from = "new_varnames", values_from = "value")

  # Split by survey data type and save separately
  df_by_survey_type<- list()
  

    # Income only, including extrapolated years
    df_by_survey_type[["inc_extrap"]]<- df %>%
      filter(datatype == "income")

    # Income only, excluding extrapolated years
    df_by_survey_type[["inc_no_extrap"]]<- df %>%
      filter(IsSurveyYear == TRUE,
             datatype == "income")

    # Consumption only, including extrapolated years
    df_by_survey_type[["cons_extrap"]]<- df %>%
      filter(datatype == "consumption")
    
    # Consumption only, excluding extrapolated years
    df_by_survey_type[["cons_no_extrap"]]<- df %>%
      filter(IsSurveyYear == TRUE,
             datatype == "consumption")
    
    # Both income and consumption, including extrapolated years
    df_by_survey_type[["inc_or_cons_extrap"]]<- df
    
    # Both income and consumption, excluding extrapolated years
    df_by_survey_type[["inc_or_cons_no_extrap"]]<- df %>%
      filter(IsSurveyYear == TRUE)
    
    
  # Get a vector of names of the different survey types
  survey_types<- names(df_by_survey_type)
  
  # save the separated csvs
  for(survey_type in survey_types){
    
  filename<- paste0(survey_type,".csv")
  fp<- paste0("data/", filename)
  
  write.csv(df_by_survey_type[[survey_type]], fp, row.names = FALSE)
  
  }
  




```


## Build the controls in google sheet

```{r}

gsheets_ids<- c(
  "1K572GXWWoGfFfKU7mGRxv7ghgs1p1onhHm-pIViNDeE", #across the distribution
  "1-65yg7odNc6wym6VxXc5Sj0Dt0A9QPaL8mfzUhc1yHE" #poverty
  )

gsheets_ids<- c(
  "1-65yg7odNc6wym6VxXc5Sj0Dt0A9QPaL8mfzUhc1yHE" #poverty
  )

for(gsheets_id in gsheets_ids){

build_OWID_controls(gsheets_id)
  
}
   
```

