---
title: "Original and adjusted global distributions"
author: "Joe Hasell"
date: "25/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data prepared by Shahid.
```{r}
df<- read.csv("full_distribution.csv")

```




Rearranging slightly, this is stored in the following format:
```{r}

df<- df %>%
  pivot_longer(
    cols = !c(CountryName, RequestYear),
    names_to = "percentile",
    names_prefix = "P",
    names_transform = list(percentile = as.numeric),
    values_to = "daily_income_or_consumption"
  )

head(df)
```

I'm not totally sure how Shahid calculated this, and hence not exactly sure what the percentile '0.0' corresponds to.

In any case, it looks like there's some kinky behaviour at this 0.0 value.

Moreover, steps of 1/10th of a percentile is too high a resolution: it results in e.g. the income threshold of P10.1 and P10.2 being exactly the same (to 2 d.p.).

Let's only keep whole-number percentiles (i.e. P1, P2... P99).


```{r}

perc_keep<- seq(1,99,1) # The set of fractiles to keep

df<- df %>%
  filter(percentile %in% perc_keep)
```