---
title: "Prepare interpolated gross and net top incomes data set"
author: "Joe Hasell"
date: "04/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Country names and regions

This code block takes the WID country-region mapping file and transforms it to a more helpful shape, and then exports a csv that I have then run through the OWID country-standardizer tool.

```{r}
# df_regions<- read.csv("original/wid-regions-list.csv")
# 
# 
# # pivot longer
# df_regions_long <- df_regions %>%
#   pivot_longer(
#     cols = !c(region, year),
#     names_to = "country",
#     values_to = "region_dummy"
#   )
# 
# # keep only where dummy=1 (in order to merge region into other datasets)
# df_regions_long <- df_regions_long %>%
#   filter(region_dummy==1)
# 
# 
# write.csv(df_regions_long, "wid_country_regions.csv", row.names = FALSE)


```

Grab top 10% share of pretax income
```{r}
wid_pretax_sh<- read.csv("original/WID pre-tax shares.csv")

wid_pretax_sh_top10<-  wid_pretax_sh %>%
  filter(percentile == "p90p100") %>%
  select(country, year, value) %>%
  rename(pre_tax_top10 = value,
         countryCode = country)

```


Grab top 10% share of post_tax income
```{r}
wid_posttax_sh<- read.csv("original/WID post-tax shares.csv")

wid_posttax_sh_top10<-  wid_posttax_sh %>%
  filter(percentile == "p90p100") %>%
  select(country, year, value) %>%
  rename(post_tax_top10 = value,
         countryCode = country)

```

Grab WID region classification and mapping to OWID country names
```{r}
country_region<- read.csv("original/WID regions_country_standardized.csv") %>%
  rename(countryCode = WID.code)
```

Merge pre- and post-tax top shares and WID region classification
```{r}

wid_sh_top10<- left_join(wid_pretax_sh_top10, wid_posttax_sh_top10)

wid_sh_top10<- left_join(wid_sh_top10, country_region)

wid_sh_top10<- wid_sh_top10 %>%
  select(-country) %>%
  rename(country = Our.World.In.Data.Name)


```

```{r}
# Export clean csv

write.csv(wid_sh_top10, "clean/WID_top10_shares.csv", row.names = FALSE)

```

