---
title: "Maternal Mortality Rate - Gapminder & WHO"
output: html_document
---

Maternal mortality ratio...

```{r, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
install.packages('janitor',repos = "http://cran.us.r-project.org")
```

Load libraries
```{r}
library(readxl)
library(dplyr)
library(janitor)
library(tidyr)
library(ggplot2)
date <- Sys.Date()

```

Download data from Gapminder and save it locally
```{r}
dir.create("data/input", recursive = TRUE)
download.file('https://www.gapminder.org/documentation/documentation/gapdata010.xls', paste0('data/input/',date,'_gapminder_data.xls'))


```

Read in the Gapminder data
```{r}

df <- read_xls(paste0('data/input/',date,'_gapminder_data.xls'), col_names = TRUE)
df <- df[-(1:16),]
```
The first three year ranges seem to be wrong as in another source (https://docs.google.com/spreadsheets/u/0/d/14ZtQy9kd0pMRKWg_zKsTg3qKHoGtflj-Ekal9gIPZ4A/pub?gid=1#) they cover only 10 years rather than 110. I will replace them manually here with the middle of each decade, as is used in the other data from the same source.

There are also two cases where the year for Finland has been entered wrongly three times, there are two 1772s, 1775s & 1967s, the second of each should be 1872, 1875 and 1967 respectively.

There are also two errors for New Zealand, and one error for both Sweden and the US.
```{r}

df$year[1:3] <- c(1875, 1885,1895)
df['year'][df$year == '1772' & df$Country == 'Finland' & df$`Maternal deaths`== 487,] <- '1872'
df['year'][df$year == '1775' & df$Country == 'Finland' & df$`Maternal deaths` == 629,] <- '1875'
df['year'][df$year == '1967' & df$Country == 'Finland' & df$`Maternal deaths` == 77,] <- '1957'
df['year'][df$year == '1967' & df$Country == 'Sweden' & df$`Maternal deaths` == 39,] <- '1957'
df['year'][df$year == '1967' & df$Country == 'United States' & df$`Maternal deaths` == 1766.28,] <- '1957'
df['year'][df$year =='1989-02'& df$Country == 'New Zealand',] <- '1889-02'

#Sri Lankas MMR drops to 0 in 1990 which seems wrong - we set it to NA instead.
df['MMR'][df$Country == "Sri Lanka" & df$year == 1990,] <- NA
#Dropping an erroneous row for NZ in 1950 - there are two values but the first is incorrect. You can using the maternal deaths and live births columns. 
df <- df[-which(df$year == '1950' & df$Country == 'New Zealand' & df$MMR == 90),]


```



For some rows a range of years is given for a particular maternal mortality rate. We use this function to find the mid-value of years given.
```{r}

mid_year <- function(year) {
  if(grepl("-", year)){
    year_split <-  unlist(strsplit(year, split = "-"))
    year_begin <- year_split[1]
    year_end_len <- nchar(year_split[2])
    year_end_pref <- 4 - year_end_len
    
    if (year_end_pref > 0) {
      year_pref <- substr(year_begin, 1, year_end_pref)
      year_end <- as.numeric(paste(c(year_pref, year_split[2]),collapse = ""))
      year_out <- round(mean(c(as.numeric(year_begin), year_end)))
    }else{
      year_out <- round(mean(as.numeric(c(year_begin,year_split[2]))))
    }
  }else{
    year_out <- year
  }
  
  return(year_out)
}
year_out_all <- unlist(lapply(df$year, mid_year))
df$mid_year <- as.numeric(year_out_all)
```

Let's check the mid-year estimates have worked as expected:
```{r}
df %>% select(year, mid_year) %>% filter(grepl("-",year)) %>% distinct()

```

Keeping only the columns we want to take forward.
```{r}
df <- df %>% 
  select(entity = Country, year = mid_year, maternal_mortality_rate = MMR)

```

Read in data from WDI - in future iterations we will read this in from ETL so it is auto updated. The WDI sources this variable from the WHO so we will henceforth refer to it as the WHO data.

```{r}
download.file('https://api.worldbank.org/v2/en/indicator/SH.STA.MMRT?downloadformat=csv', paste0('data/input/',date,'_wdi_maternal_mortality.zip'))

unzip(paste0('data/input/',date,'_wdi_maternal_mortality.zip'), exdir = "data/input")
des_file <- unzip(paste0('data/input/',date,'_wdi_maternal_mortality.zip'), list=TRUE)
file_name <- des_file$Name[startsWith( des_file$Name,"API_SH.STA.MMRT_DS2")]

who <- clean_names(read.csv(paste0('data/input/', file_name), skip = 4))

```

Pivot and clean the WHO data to long format
```{r}

who_clean <- who %>% 
  select(country_name, starts_with("X")) %>% 
  select(-x) %>% 
  pivot_longer(starts_with("x")) %>% 
  mutate(year = as.numeric(gsub("x", "", name))) %>% 
  select(entity = country_name, year, maternal_mortality_rate = value)



```

Check the Gapminder countries are in the WHO dataset
```{r}

unique(df$entity) %in% who_clean$entity


```

Combine the two datasets and add a source column so we can keep track of where each datapoin comes from. 
```{r}

df$source <- 'GapMinder'
who_clean$source <- 'WHO_GHO'
df_all <- rbind(df, who_clean)

df_all$maternal_mortality_rate <- as.numeric(df_all$maternal_mortality_rate)
```


Dealing with duplicate years - there are two identical values for NZ in 1950, we can drop one with distinct(). We remove NAs where there is a non-NA value in either Gapminder or WHO.  If there are values for both Gapminder and WHO GHO we have a preference for WHO GHO. 

```{r}
df_check <-df_all %>% 
  group_by(entity, year) %>% 
  add_count() %>% 
  ungroup()


df_sel <- df_check[-which(df_check$n >1 & is.na(df_check$maternal_mortality_rate)),]

df_sel <-df_sel %>% 
  group_by(entity, year) %>% 
  add_count() %>% 
  ungroup() %>% 
  distinct()

df_sel <- df_sel[-which(df_sel$nn >1 & df_sel$source == 'GapMinder'),]

df_sel %>% 
  group_by(entity, year) %>% 
  add_count() %>% 
  filter(nnn >1) %>% 
  ungroup()


df_sel <- df_sel %>% select(-c(n, nn))
```
Standardising the country names for use on OWID. Writing the data out.

```{r}
stand_countries <- read.csv("data/input/countries_to_standardise_country_standardized.csv")

df_fin <- df_sel %>% left_join(.,stand_countries, by = c("entity" = "Country")) %>% 
  mutate(entity = Our.World.In.Data.Name) %>% 
  select(-c(Our.World.In.Data.Name, source)) 

df_fin$maternal_mortality_rate[is.na(df_fin$maternal_mortality_rate)] <- ""

```
We don't want to include all of the available entities in the data as it is difficult for us to define them clearly.

```{r}
entities_to_drop <- c("Africa Eastern and Southern", "Africa Western and Central", "Arab World", "Early-demographic dividend", "East Asia & Pacific", "East Asia & Pacific (excluding high income)", "East Asia & Pacific (IDA & IBRD countries)", "Euro area", "Europe & Central Asia (excluding high income)", "Europe & Central Asia (IDA & IBRD countries)", "Fragile and conflict affected situations", "Heavily indebted poor countries (HIPC)", "High income", "IBRD only", "IDA & IBRD total", "IDA blend", "IDA only", "IDA total", "Late-demographic dividend", "Latin America & Caribbean (excluding high income)", "Latin America & the Caribbean (IDA & IBRD countries)", "Least developed countries: UN classification", "Middle East & North Africa (excluding high income)", 
                      "Middle East & North Africa (IDA & IBRD countries)", "Middle income", "Not classified",  "Other small states", "Pacific island small states", "Post-demographic dividend", "Pre-demographic dividend", "Small states", "South Asia (IDA & IBRD)", "Sub-Saharan Africa (excluding high income)", "Sub-Saharan Africa (IDA & IBRD countries)", "Low income", "Low & middle income", "Lower middle income", "Upper income", "Upper-middle income")

df_fin <- df_fin %>% 
 filter(!entity %in% entities_to_drop)
```




Plot to see what it looks like:
```{r}

ggplot(data = df_fin %>% filter(entity== "Sweden"), aes(x = year, y = as.numeric(maternal_mortality_rate), group = entity, colour = entity))+
  geom_line()

```

```{r}

df_clean <- df_fin %>% 
  filter(year %in% c(2000, 2017)) %>% 
  filter(complete.cases(.)) %>%
  mutate(maternal_mortality_rate = as.numeric(maternal_mortality_rate)) %>% 
  group_by(entity) %>% 
    add_count() %>% 
  filter(n == 2)
  
df_wide <- df_clean %>% 
  pivot_wider(., names_from = 'year',names_glue = "{.value}_{year}", values_from = 'maternal_mortality_rate') %>% 
  mutate(change = ((maternal_mortality_rate_2017 - maternal_mortality_rate_2000)/maternal_mortality_rate_2000)*100, baseline = 0) %>% 
  select(entity, change, baseline) %>% 
  pivot_longer(cols = c("change", "baseline")) %>% 
  mutate(name = gsub("change", 2017, name), name = gsub("baseline", 2000, name)) %>% 
  mutate(year = as.numeric(name)) %>% 
  select(entity, year, relative_change_mmr = value)


```
Combining with the full time-line of MMR

```{r}

df_out <- df_wide %>% right_join(., df_fin, by = c("entity","year"))
df_out$relative_change_mmr[is.na(df_out$relative_change_mmr)] <- ""
```


Writing out the data
```{r}
write.csv(df_out, "data/output/maternal_mortality_rate_upload.csv", row.names = FALSE)

```