---
title: "main_joe_draft"
format: html
editor: visual
---

## PIP pipeline


```{python}
import pandas as pd
import numpy as np
import time

#from functions.PIP_API_query import pip_query_country, pip_query_region
from functions.standardize_entities import standardize_entities
from functions.upload import upload_to_s3

from create_dataset_functions import *
```


We ask to confirm whether 
JH Comment - Pablo, my feeling is that it would be better to run the relative_poverty.py and extract_percentiles.py scripts also from main.py, but where you ask for user input to determine whether they are run or not. 

```{python}
print('CAUTION: If a full update is needed, first you have to run relative_poverty.py and extract_percentiles.py')
print('These codes take several hours to run, that\'s why they are not available here')
question = "Do you want to continue?"
query_yes_no(question)

```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
