---
title: "Format Michalis' output for explorer"
author: "Joe Hasell"
date: "21/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RCurl)
```


```{r}
df <- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/MichalisMoatsos/MainDataCountriesOnly.csv", row.names = 1, header = T, sep = ";", dec = ",")

```

Map 3-letter codes to country names
```{r}
name_mapping <- read.csv("entity_names_three_letter_country_standardized.csv") %>%
  rename(Entity = Country)

df<- left_join(df,name_mapping)

df<- df %>%
  select(-Entity) %>%
  rename(Entity = Our.World.In.Data.Name) %>%
  select(Entity, everything())

```

Rename coumns to work with explorers
```{r}
names(df) <- gsub("[.]", "_", names(df))

```



Absolute poverty controls

```{r}


pov_lines<- data.frame(rough_pov_line =  c("1_9",
                    "3_2",
                    "5_5",
                    "10",
                    "15",
                    "20",
                    "30",
                    "40"),
                          clean_pov_line = c( "$1.90 a day",
                    "$3.20 a day",
                    "$5.50 a day",
                    "$10 a day",
                    "$15 a day",
                    "$20 a day",
                    "$30 a day",
                    "$40 a day"))



metrics<- data.frame(metric_dropdown = c("Share in poverty",
                            "Number of people in poverty",
                            "Average shortfall to poverty line, per day",
                            "Average shortfall to poverty line, per year",
                            "Total shortfall to poverty line, per day",
                            "Total shortfall to poverty line, per year",
                            "Income gap ratio",
                            "Poverty gap index"),
                     metric_slug_snippet = c("share_of_population_below_poverty_line",
                                               "total_number_of_people_below_poverty_line",
                                             "average_shortfall_per_day",
                                             "average_shortfall_per_year",
                                             "total_shortfall_per_day",
                                             "total_shortfall_per_year",
                                             "income_gap_ratio",
                                             "poverty_gap_index"),
                     title_snippet = c("Poverty: Share of population living on less than",
                                       "Poverty: Number of people living on less than",
                                       "Depth of poverty: Average daily shortfall to",
                                       "Depth of poverty: Average annual shortfall to",
                                       "Total daily shortfall to a poverty line of",
                                       "Total annual shortfall to a poverty line of",
                                       "Depth of poverty: Income gap ratio at",
                                       "Poverty gap index at"))



abs_pov_controls<- expand.grid("metric_dropdown" = metrics$metric_dropdown,
                               "rough_pov_line" = pov_lines$rough_pov_line) 

abs_pov_controls <- left_join(abs_pov_controls, pov_lines) %>%
  rename(pov_line_dropdown = clean_pov_line)

abs_pov_controls <- left_join(abs_pov_controls, metrics) %>%
  mutate(ySlugs = paste0("X_", rough_pov_line, "_per_day___", metric_slug_snippet),
         title = paste0(title_snippet, " ", pov_line_dropdown))


```

Relative poverty headcount controls
```{r}


pov_lines<- data.frame(rough_pov_line =  c("X_40__of_median_income",
                    "X_50__of_median_income",
                    "X_60__of_median_income"),
                          clean_pov_line = c("40% of median",
                    "50% of median",
                    "60% of median"))



metrics<- data.frame(metric_dropdown = c("Share in poverty",
                            "Number of people in poverty",
                            "Average shortfall to poverty line, per day",
                            "Average shortfall to poverty line, per year",
                            "Total shortfall to poverty line, per day",
                            "Total shortfall to poverty line, per year",
                            "Income gap ratio",
                            "Poverty gap index"),
                     metric_slug_snippet = c("share_of_population_below_poverty_line",
                                               "total_number_of_people_below_poverty_line",
                                             "average_shortfall_per_day",
                                             "average_shortfall_per_year",
                                             "total_shortfall_per_day",
                                             "total_shortfall_per_year",
                                             "income_gap_ratio",
                                             "poverty_gap_index"),
            title_snippet = c("Relative poverty: Share of population living on less than",
                               "Relative poverty: Number of people living on less than",
                              "Depth of relative poverty: Average daily shortfall to",
                              "Depth of relative poverty: Average annual shortfall to",
                              "Total daily shortfall to a poverty line of",
                              "Total annual shortfall to a poverty line of",
                              "Depth of relative poverty: Income gap ratio at",
                              "Poverty gap index at"))



rel_pov_controls<- expand.grid("metric_dropdown" = metrics$metric_dropdown,
                               "rough_pov_line" = pov_lines$rough_pov_line) 

rel_pov_controls <- left_join(rel_pov_controls, pov_lines) %>%
  mutate(pov_line_dropdown = paste0("Relative line: ", clean_pov_line))

rel_pov_controls <- left_join(rel_pov_controls, metrics) %>%
  mutate(ySlugs = paste0(rough_pov_line, "___", metric_slug_snippet),
         title = paste0(title_snippet, " ", clean_pov_line))


```

Join controls for the different headcount controls
```{r}
poverty_controls<- bind_rows(abs_pov_controls, rel_pov_controls) %>%
  select(title, metric_dropdown, pov_line_dropdown, ySlugs)

```


Across the distribution controls



Inequality controls
```{r}


metric_dropdown_non_pov<- c("Mean income or consumption per day",
                            "Median income or consumption per day",
                            "P10: the income or consumption marking the poorest tenth",
                            "P90: the income or consumption marking the richest tenth",
                            "P90:P10 ratio",
                            "Gini coefficient")

```

Join different control types together
```{r}
controls<- bind_rows(poverty_controls)
  
  
```

Rework controls for each survey data type
```{r}

controls_inc_only<- controls %>%
  mutate(title = gsub("of median", "of median income", title),
         survey_type_dropdown = "Income only",
         extrapolate_checkbox = "'false",
         tableSlug = "Income_only")


controls_cons_only<- controls %>%
  mutate(title = gsub("of median", "of median consumption", title),
         survey_type_dropdown = "Consumption only",
         extrapolate_checkbox = "'false",
         tableSlug = "Consumption_only")

controls_inc_and_cons_no_extrap<- controls %>%
  mutate(title = gsub("of median", "of median income or consumption", title),
         survey_type_dropdown = "Either income or consumption",
         extrapolate_checkbox = "'false",
         tableSlug = "Inc_or_cons_no_extrap")


controls_inc_and_cons_plus_extrap<- controls %>%
  mutate(title = gsub("of median", "of median income or consumption", title),
         survey_type_dropdown = "Either income or consumption",
         extrapolate_checkbox = "'true",
         tableSlug = "Inc_or_cons_plus_extrap")


```


Join survey type controls together
```{r}
all_types_controls<- bind_rows(controls_inc_and_cons_plus_extrap, controls_inc_and_cons_no_extrap, controls_inc_only, controls_cons_only )

all_types_controls<- all_types_controls %>%
  select(title, 
         metric_dropdown, 
         pov_line_dropdown, 
         survey_type_dropdown, 
         extrapolate_checkbox, 
         tableSlug, 
         ySlugs)

all_types_controls<- all_types_controls %>%
  mutate(yAxisMin = 0)

names(all_types_controls)<- c("title",
                              "Metric Dropdown", 
                              "Poverty line Dropdown", 
                              "Household survey data type Radio", 
                              "Show extrapolated estimates for years with no survey data Checkbox", 
                              "tableSlug", 
                              "ySlugs",
                              "yAxisMin")


```



Write explorer controls to csv
```{r}
write.csv(all_types_controls, "explorer_controls.csv", row.names = F)
```


Split dfs by survey type

```{r}

income_only<- df %>%
  filter(IsSurveyYear == TRUE,
         datatype == "income")

write.csv(income_only, "income_only.csv", row.names = FALSE)
  

consumption_only<- df %>%
  filter(IsSurveyYear == TRUE,
         datatype == "consumption")

write.csv(consumption_only, "consumption_only.csv", row.names = FALSE)
  


inc_and_cons_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE)

write.csv(inc_and_cons_no_extrap, "inc_and_cons_no_extrap.csv", row.names = FALSE)
  
write.csv(df, "inc_and_cons_plus_extrap.csv", row.names = FALSE)




```

