---
title: "Format Michalis' output for explorer"
author: "Joe Hasell"
date: "21/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RCurl)
```


```{r}
df <- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/MichalisMoatsos/MainDataCountriesOnly.csv", row.names = 1, header = T, sep = ";", dec = ",")

```

Map 3-letter codes to country names
```{r}
name_mapping <- read.csv("entity_names_three_letter_country_standardized.csv") %>%
  rename(Entity = Country)

df<- left_join(df,name_mapping)

df<- df %>%
  select(-Entity) %>%
  rename(Entity = Our.World.In.Data.Name) %>%
  select(Entity, everything())

```

Separate out Urban/Rural entities (except for Argentina, where we only have Urban-only data. For Argentina we will add an annotation. )
```{r}


```


Rename columns to work with explorers
```{r}
names(df) <- gsub("[.]", "_", names(df))

```
## Controls and data for poverty explorer
```{r}
poverty_metrics<- data.frame(metric_dropdown = c("Share in poverty",
                            "Number of people in poverty",
                            "Average shortfall to poverty line, per day",
                            "Average shortfall to poverty line, per year",
                            "Total shortfall to poverty line, per day",
                            "Total shortfall to poverty line, per year",
                            "Income gap ratio",
                            "Poverty gap index"),
                     metric_slug_snippet = c("share_of_population_below_poverty_line",
                                               "total_number_of_people_below_poverty_line",
                                             "average_shortfall_per_day",
                                             "average_shortfall_per_year",
                                             "absolute_poverty_gap_per_day",
                                             "absolute_poverty_gap",
                                             "income_gap_ratio",
                                             "poverty_gap_index"))
```

Absolute poverty controls

```{r}


pov_lines<- data.frame(rough_pov_line =  c("1_9",
                    "3_2",
                    "5_5",
                    "10",
                    "15",
                    "20",
                    "30",
                    "40"),
                          clean_pov_line = c( "$1.90 a day",
                    "$3.20 a day",
                    "$5.50 a day",
                    "$10 a day",
                    "$15 a day",
                    "$20 a day",
                    "$30 a day",
                    "$40 a day"))



poverty_metrics<- poverty_metrics %>%
  mutate(title_snippet = c("Poverty: Share of population living on less than",
                           "Poverty: Number of people living on less than",
                            "Depth of poverty: Average daily shortfall below",
                            "Depth of poverty: Average annual shortfall below",
                            "Total daily shortfall to a poverty line of",
                            "Total annual shortfall below a poverty line of",
                            "Depth of poverty: Income gap ratio at",
                            "Poverty gap index at"))



abs_pov_controls<- expand.grid("metric_dropdown" = poverty_metrics$metric_dropdown,
                               "rough_pov_line" = pov_lines$rough_pov_line) 

abs_pov_controls <- left_join(abs_pov_controls, pov_lines) %>%
  rename(pov_line_dropdown = clean_pov_line)

abs_pov_controls <- left_join(abs_pov_controls, poverty_metrics) %>%
  mutate(ySlugs = paste0("X_", rough_pov_line, "_per_day___", metric_slug_snippet),
         title = paste0(title_snippet, " ", pov_line_dropdown))


```

Relative poverty headcount controls
```{r}


pov_lines<- data.frame(rough_pov_line =  c("X_40__of_median_income",
                    "X_50__of_median_income",
                    "X_60__of_median_income"),
                          clean_pov_line = c("40% of median",
                    "50% of median",
                    "60% of median"))


poverty_metrics<- poverty_metrics %>%
  mutate(title_snippet = c("Relative poverty: Share of population living on less than",
                           "Relative poverty: Number of people living on less than",
                           "Depth of relative poverty: Average daily shortfall to",
                           "Depth of relative poverty: Average annual shortfall to",
                           "Total daily shortfall to a poverty line of",
                           "Total annual shortfall to a poverty line of",
                           "Depth of relative poverty: Income gap ratio at",
                           "Poverty gap index at"))



rel_pov_controls<- expand.grid("metric_dropdown" = poverty_metrics$metric_dropdown,
                               "rough_pov_line" = pov_lines$rough_pov_line) 

rel_pov_controls <- left_join(rel_pov_controls, pov_lines) %>%
  mutate(pov_line_dropdown = paste0("Relative line: ", clean_pov_line))

rel_pov_controls <- left_join(rel_pov_controls, poverty_metrics) %>%
  mutate(ySlugs = paste0(rough_pov_line, "___", metric_slug_snippet),
         title = paste0(title_snippet, " ", clean_pov_line))


```

Join controls for the different headcount controls
```{r}
poverty_controls<- bind_rows(abs_pov_controls, rel_pov_controls) %>%
  select(title, metric_dropdown, pov_line_dropdown, ySlugs)

```


Function to repeat controls for each survey data type
```{r}


repeat_controls_by_survey_type<- function(controls_df){

controls_by_survey_type<- list()

controls_by_survey_type[["controls_inc_or_cons"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Either income or consumption",
         extrapolate_checkbox = "'true",
         tableSlug = "inc_or_cons")


controls_by_survey_type[["controls_inc_or_cons_no_extrap"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Either income or consumption",
         extrapolate_checkbox = "'false",
         tableSlug = "inc_or_cons_no_extrap")


controls_by_survey_type[["controls_inc_only"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Income only",
         extrapolate_checkbox = "'true",
         tableSlug = "inc")

controls_by_survey_type[["controls_inc_only_no_extrap"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Income only",
         extrapolate_checkbox = "'false",
         tableSlug = "inc_no_extrap")



controls_by_survey_type[["controls_cons_only"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Consumption only",
         extrapolate_checkbox = "'true",
         tableSlug = "cons")


controls_by_survey_type[["controls_cons_only_no_extrap"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Consumption only",
         extrapolate_checkbox = "'false",
         tableSlug = "cons_no_extrap")




return(controls_by_survey_type)

}

```

Repeat controls for poverty
```{r}
poverty_controls_by_survey_type<- repeat_controls_by_survey_type(poverty_controls)
```


Join survey type controls together
```{r}
poverty_controls_all_survey_types<- bind_rows(poverty_controls_by_survey_type)

poverty_controls_all_survey_types<- poverty_controls_all_survey_types %>%
  select(title, 
         metric_dropdown, 
         pov_line_dropdown, 
         survey_type_dropdown, 
         extrapolate_checkbox, 
         tableSlug, 
         ySlugs)

poverty_controls_all_survey_types<- poverty_controls_all_survey_types %>%
  mutate(yAxisMin = 0)

names(poverty_controls_all_survey_types)<- c("title",
                              "Metric Dropdown", 
                              "Poverty line Dropdown", 
                              "Household survey data type Radio", 
                              "Show extrapolated estimates for years with no survey data Checkbox", 
                              "tableSlug", 
                              "ySlugs",
                              "yAxisMin")


```




Write poverty explorer controls to csv
```{r}
write.csv(poverty_controls_all_survey_types, "poverty_explorer_controls.csv", row.names = F)
```



## Incomes across the distribution explorer controls

```{r}

# thresholds<- c("P10 (poorest tenth)",
#               "P20",
#               "P30",
#               "P40",
#               "Median – P50",
#               "P60",
#               "P70",
#               "P80",
#               "P90 (richest tenth)",
#               clean<- "1","2","3","4","5",
#                         "6", "7", "8", "9")
# 
# deciles<- c("Decile 1 (poorest tenth)",
#             "Decile 2",
#             "Decile 3",
#             "Decile 4",
#             "Decile 5",
#             "Decile 6",
#             "Decile 7",
#             "Decile 8",
#             "Decile 9",
#             "Decile 10 (richest tenth)")
# 

```


## Controls for Inequality explorer
```{r}

```





##Split dfs by survey type



Add annotation for survey type (where this changes over time, say 'Income data before X, consumption after')
```{r}


```




```{r}

# Income only, including extrapolated years
inc<- df %>%
  filter(datatype == "income")

write.csv(inc, "data/inc.csv", row.names = FALSE)

# Income only, excluding extrapolated years
inc_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE,
         datatype == "income")

write.csv(inc_no_extrap, "data/inc_no_extrap.csv", row.names = FALSE)

  
# Consumption only, including extrapolated years
cons<- df %>%
  filter(datatype == "consumption")

write.csv(cons, "data/cons.csv", row.names = FALSE)

# Consumption only, excluding extrapolated years
cons_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE,
         datatype == "consumption")  

write.csv(cons_no_extrap, "data/cons_no_extrap.csv", row.names = FALSE)


# Both income and consumption, including extrapolated years
write.csv(df, "data/inc_or_cons.csv", row.names = FALSE)


# Both income and consumption, excluding extrapolated years
inc_or_cons_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE)

# Check for years where there are both income and consumption surveys

write.csv(inc_or_cons_no_extrap, "data/inc_or_cons_no_extrap.csv", row.names = FALSE)
  




```



