---
title: "Format Michalis' output for explorer"
author: "Joe Hasell"
date: "21/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RCurl)
library(knitr)
library(readxl)
library(openxlsx)
```

Read in Michalis' povcal data
```{r}
df <- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/MichalisMoatsos/PovcalNetMainDataCountriesOnly.csv", row.names = 1, header = T, sep = ";", dec = ",")

```

Map 3-letter codes to country names
```{r}
name_mapping <- read.csv("entity_names_three_letter_country_standardized.csv") %>%
  rename(Entity = Country)

df<- left_join(df,name_mapping)

df<- df %>%
  select(-Entity) %>%
  rename(Entity = Our.World.In.Data.Name) %>%
  select(Entity, everything())

```

Rename vars (We should do rearrange Michalis' script to structure the variables like this and then delete this section)

```{r}


orig_varnames<- c(

"Entity",
"Year",
"X.1.per.day...share.of.population.below.poverty.line",
"X.1.per.day...poverty.gap.index",
"X.1.per.day...total.number.of.people.below.poverty.line",
"X.1.per.day...absolute.poverty.gap",
"X.1.per.day...income.gap.ratio",
"X.1.per.day...watts.index",
"X.1.9.per.day...share.of.population.below.poverty.line",
"X.1.9.per.day...poverty.gap.index",
"X.1.9.per.day...total.number.of.people.below.poverty.line",
"X.1.9.per.day...absolute.poverty.gap",
"X.1.9.per.day...income.gap.ratio",
"X.1.9.per.day...watts.index",
"X.3.2.per.day...share.of.population.below.poverty.line",
"X.3.2.per.day...poverty.gap.index",
"X.3.2.per.day...total.number.of.people.below.poverty.line",
"X.3.2.per.day...absolute.poverty.gap",
"X.3.2.per.day...income.gap.ratio",
"X.3.2.per.day...watts.index",
"X.5.5.per.day...share.of.population.below.poverty.line",
"X.5.5.per.day...poverty.gap.index",
"X.5.5.per.day...total.number.of.people.below.poverty.line",
"X.5.5.per.day...absolute.poverty.gap",
"X.5.5.per.day...income.gap.ratio",
"X.5.5.per.day...watts.index",
"X.10.per.day...share.of.population.below.poverty.line",
"X.10.per.day...poverty.gap.index",
"X.10.per.day...total.number.of.people.below.poverty.line",
"X.10.per.day...absolute.poverty.gap",
"X.10.per.day...income.gap.ratio",
"X.10.per.day...watts.index",
"X.15.per.day...share.of.population.below.poverty.line",
"X.15.per.day...poverty.gap.index",
"X.15.per.day...total.number.of.people.below.poverty.line",
"X.15.per.day...absolute.poverty.gap",
"X.15.per.day...income.gap.ratio",
"X.15.per.day...watts.index",
"X.20.per.day...share.of.population.below.poverty.line",
"X.20.per.day...poverty.gap.index",
"X.20.per.day...total.number.of.people.below.poverty.line",
"X.20.per.day...absolute.poverty.gap",
"X.20.per.day...income.gap.ratio",
"X.20.per.day...watts.index",
"X.30.per.day...share.of.population.below.poverty.line",
"X.30.per.day...poverty.gap.index",
"X.30.per.day...total.number.of.people.below.poverty.line",
"X.30.per.day...absolute.poverty.gap",
"X.30.per.day...income.gap.ratio",
"X.30.per.day...watts.index",
"X.40.per.day...share.of.population.below.poverty.line",
"X.40.per.day...poverty.gap.index",
"X.40.per.day...total.number.of.people.below.poverty.line",
"X.40.per.day...absolute.poverty.gap",
"X.40.per.day...income.gap.ratio",
"X.40.per.day...watts.index",
"X40..of.median.income...share.of.population.below.poverty.line",
"X40..of.median.income...poverty.gap.index",
"X40..of.median.income...total.number.of.people.below.poverty.line",
"X40..of.median.income...absolute.poverty.gap",
"X40..of.median.income...income.gap.ratio",
"X40..of.median.income...watts.index",
"X50..of.median.income...share.of.population.below.poverty.line",
"X50..of.median.income...poverty.gap.index",
"X50..of.median.income...total.number.of.people.below.poverty.line",
"X50..of.median.income...absolute.poverty.gap",
"X50..of.median.income...income.gap.ratio",
"X50..of.median.income...watts.index",
"X60..of.median.income...share.of.population.below.poverty.line",
"X60..of.median.income...poverty.gap.index",
"X60..of.median.income...total.number.of.people.below.poverty.line",
"X60..of.median.income...absolute.poverty.gap",
"X60..of.median.income...income.gap.ratio",
"X60..of.median.income...watts.index",
"MeanPerMonth",
"MeanPerMonth_estimated",
"MedianPerMonth",
"MedianPerMonth_estimated",
"PPP",
"Population",
"Decile.1...share.of.income.or.consumption",
"Decile.2...share.of.income.or.consumption",
"Decile.3...share.of.income.or.consumption",
"Decile.4...share.of.income.or.consumption",
"Decile.5...share.of.income.or.consumption",
"Decile.6...share.of.income.or.consumption",
"Decile.7...share.of.income.or.consumption",
"Decile.8...share.of.income.or.consumption",
"Decile.9...share.of.income.or.consumption",
"Decile.10...share.of.income.or.consumption",
"Decile.1...threshold.of.income.or.consumption.per.month",
"Decile.2...threshold.of.income.or.consumption.per.month",
"Decile.3...threshold.of.income.or.consumption.per.month",
"Decile.4...threshold.of.income.or.consumption.per.month",
"Decile.5...threshold.of.income.or.consumption.per.month",
"Decile.6...threshold.of.income.or.consumption.per.month",
"Decile.7...threshold.of.income.or.consumption.per.month",
"Decile.8...threshold.of.income.or.consumption.per.month",
"Decile.9...threshold.of.income.or.consumption.per.month",
"Decile.1...average.income.or.consumption.per.month",
"Decile.2...average.income.or.consumption.per.month",
"Decile.3...average.income.or.consumption.per.month",
"Decile.4...average.income.or.consumption.per.month",
"Decile.5...average.income.or.consumption.per.month",
"Decile.6...average.income.or.consumption.per.month",
"Decile.7...average.income.or.consumption.per.month",
"Decile.8...average.income.or.consumption.per.month",
"Decile.9...average.income.or.consumption.per.month",
"Decile.10...average.income.or.consumption.per.month",
"Gini.index",
"Gini_estimated",
"Polarization",
"Polarization_estimated",
"MLD",
"MLD_estimated",
"Palma",
"P90.P10.ratio",
"P90.50.ratio",
"Entropy_0_5",
"Entropy_1_0",
"Entropy_1_5",
"Entropy_2_0",
"Atkinson_0_5",
"Atkinson_1_0",
"Atkinson_1_5",
"Atkinson_2_0",
"Theil_0_0",
"Theil_0_5",
"Theil_1_0",
"Theil_1_5",
"Theil_2_0",
"Var.Coeff",
"datayear",
"isinterpolated",
"usemicrodata",
"coveragetype",
"datatype",
"IsSurveyYear",
"OriginalMedian",
"OriginalMean",
"OriginalDecileShares",
"MonotonicityBreaks",
"RowsWithIncreasingHeadcount",
"DataframeRowsForGpinter",
"LessThan33Rows",
"ISO3DataYearCovType",
"ISO3Year"
)


new_varnames<- c(

"Entity",
"Year",
"headcount_ratio_1_00",
"poverty_gap_index_1_00",
"headcount_1_00",
"total_shortall_annual_1_00",
"income_gap_ratio_1_00",
"watts_index_1_00",
"headcount_ratio_1_90",
"poverty_gap_index_1_90",
"headcount_1_90",
"total_shortall_annual_1_90",
"income_gap_ratio_1_90",
"watts_index_1_90",
"headcount_ratio_3_20",
"poverty_gap_index_3_20",
"headcount_3_20",
"total_shortall_annual_3_20",
"income_gap_ratio_3_20",
"watts_index_3_20",
"headcount_ratio_5_50",
"poverty_gap_index_5_50",
"headcount_5_50",
"total_shortall_annual_5_50",
"income_gap_ratio_5_50",
"watts_index_5_50",
"headcount_ratio_10_00",
"poverty_gap_index_10_00",
"headcount_10_00",
"total_shortall_annual_10_00",
"income_gap_ratio_10_00",
"watts_index_10_00",
"headcount_ratio_",
"poverty_gap_index_15_00",
"headcount_15_00",
"total_shortall_annual_15_00",
"income_gap_ratio_15_00",
"watts_index_15_00",
"headcount_ratio_20_00",
"poverty_gap_index_20_00",
"headcount_20_00",
"total_shortall_annual_20_00",
"income_gap_ratio_20_00",
"watts_index_20_00",
"headcount_ratio_30_00",
"poverty_gap_index_30_00",
"headcount_30_00",
"total_shortall_annual_30_00",
"income_gap_ratio_30_00",
"watts_index_30_00",
"headcount_ratio_40_00",
"poverty_gap_index_40_00",
"headcount_40_00",
"total_shortall_annual_40_00",
"income_gap_ratio_40_00",
"watts_index_40_00",
"headcount_ratio_40_median",
"poverty_gap_index_40_median",
"headcount_40_median",
"total_shortall_annual_40_median",
"income_gap_ratio_40_median",
"watts_index_40_median",
"headcount_ratio_50_median",
"poverty_gap_index_50_median",
"headcount_50_median",
"total_shortall_annual_50_median",
"income_gap_ratio_50_median",
"watts_index_50_median",
"headcount_ratio_60_median",
"poverty_gap_index_60_median",
"headcount_60_median",
"total_shortall_annual_60_median",
"income_gap_ratio_60_median",
"watts_index_60_median",
"MeanPerMonth",
"MeanPerMonth_estimated",
"MedianPerMonth",
"MedianPerMonth_estimated",
"PPP",
"Population",
"share_decile_1",
"share_decile_2",
"share_decile_3",
"share_decile_4",
"share_decile_5",
"share_decile_6",
"share_decile_7",
"share_decile_8",
"share_decile_9",
"share_decile_10",
"threshold_decile_1",
"threshold_decile_2",
"threshold_decile_3",
"threshold_decile_4",
"threshold_decile_5",
"threshold_decile_6",
"threshold_decile_7",
"threshold_decile_8",
"threshold_decile_9",
"average_decile_1",
"average_decile_2",
"average_decile_3",
"average_decile_4",
"average_decile_5",
"average_decile_6",
"average_decile_7",
"average_decile_8",
"average_decile_9",
"average_decile_10",
"Gini",
"Gini_estimated",
"Polarization",
"Polarization_estimated",
"MLD",
"MLD_estimated",
"Palma",
"P90_P10_ratio",
"P90_50_ratio",
"Entropy_0_5",
"Entropy_1_0",
"Entropy_1_5",
"Entropy_2_0",
"Atkinson_0_5",
"Atkinson_1_0",
"Atkinson_1_5",
"Atkinson_2_0",
"Theil_0_0",
"Theil_0_5",
"Theil_1_0",
"Theil_1_5",
"Theil_2_0",
"Var.Coeff",
"datayear",
"isinterpolated",
"usemicrodata",
"coveragetype",
"datatype",
"IsSurveyYear",
"OriginalMedian",
"OriginalMean",
"OriginalDecileShares",
"MonotonicityBreaks",
"RowsWithIncreasingHeadcount",
"DataframeRowsForGpinter",
"LessThan33Rows",
"ISO3DataYearCovType",
"ISO3Year"
)

varname_map<- data.frame(orig_varnames,new_varnames)


```


Swap in new varnames
```{r}
df<- df %>%
  pivot_longer(cols = !c(Entity, Year, coveragetype, datatype, ISO3Year, ISO3DataYearCovType), names_to = "orig_varnames", values_to = "value")

# Merge in new var names

df<- left_join(df, varname_map)

# check there are no unmatched vars
br<- df %>%
  filter(is.na(new_varnames))

```




Demark Urban/Rural entities.
```{r}
df<- df %>%
  mutate(Entity = if_else(coveragetype == "R",
                          paste0(Entity, " -  Rural"),
                          Entity)) %>%
  mutate(Entity = if_else(coveragetype == "U",
                          paste0(Entity, " -  Urban"),
                          Entity))

```



## Controls

Pull in the control stubs
```{r}

xlsx_url <- "https://raw.githubusercontent.com/owid/notebooks/main/JoeHasell/Copy_of_Michalis_work/control_stubs.xlsx"

control_stubs_sheet_name<- "control_stubs"

# Load the stubs as a dataframe
control_stubs<- read.xlsx(xlsx_url, sheet = control_stubs_sheet_name)

```

Make a vector of the names of the other sheets (these are the Aux sheets, which will be used to multiply-up the control stubs, replacing the various combination of values contained in the Aux sheets)
```{r}

wb<- loadWorkbook(xlsx_url)

# a vector of the names of the other sheets (the Aux vars)
aux_sheets<- names(wb)
aux_sheets<- aux_sheets[aux_sheets != "control_stubs"]

```


The first step in multiplying up the control stubs is to convert the dataframe of control stubs to a list of rows.
```{r}

running_list<- split(control_stubs, seq(nrow(control_stubs)))

```


###Functions to multiply up the control stubs with the Aux values

This function takes a dataframe (`df_block`) and within it replaces a set of key words for a set of values. The key words and replacement values are specified by `sheet_name` (string) and `replacements_row` (a 1 row dataframe). The key words are strings, constructed as "[`sheet_name`$`column_header`]", where `column_header` is each of the variable names in `replacements_row`. The replacement values are the corresponding values of `replacements_row`.

```{r}
myReplaceFun<- function(df_block, sheet_name, replacements_row){
  
    # Backslashes for regex special character escapes
  search_for_strs<- paste0("\\[", sheet_name, "\\$", names(replacements_row), "\\]")
  
  replace_with_strs<- as.vector(t(replacements_row))

  replace_vector_over_col <- function(x) (str_replace_all(x, setNames(replace_with_strs,search_for_strs)))
  
  df_block_replaced<- df_block %>% 
      mutate(across(everything(), replace_vector_over_col))

  return(df_block_replaced)
  }
    

```

Test this function
```{r}
df_block<- data.frame(col1 = c("headcount_ratio_[abs_povline$slug_suffix]","headcount_[abs_povline$slug_suffix]"), 
                      col2 = c("$[abs_povline$text] a day","$[abs_povline$text] a day"))

sheet_name<- "abs_povline"

replacements_row<- data.frame("text" = "1",
                              "slug_suffix" = "1_00",
                              "math" = "1")



kable(myReplaceFun(df_block = df_block,
                            sheet_name = sheet_name,
                            replacements_row = replacements_row))

```

A function that splits a dataframe (`df_replacements_rows`, where each row is a replacements_row in myReplaceFun) into a list of dataframes by row, applies the myReplaceFun on each of those rows, and then aggregates the resulting list back into a single dataframe. The arguments `df_block` and `sheet_name` are passed to myReplaceFun for all rows.
```{r}
applymyReplaceFunRowwise<- function(df_block, sheet_name, df_replacements_rows){

 list_replacements_rows<- split(df_replacements_rows, seq(nrow(df_replacements_rows)))
 
 list_replaced_df_blocks<- lapply(list_replacements_rows, myReplaceFun, df_block = df_block, sheet_name = sheet_name)

  df_replaced_df_blocks<- bind_rows(list_replaced_df_blocks)

  return(df_replaced_df_blocks)

}  
  
```

Test this function
```{r}

df_block<- data.frame(col1 = c("headcount_ratio_[abs_povline$slug_suffix]","headcount_[abs_povline$slug_suffix]"), 
                      col2 = c("$[abs_povline$text] a day","$[abs_povline$text] a day"))

sheet_name<- "abs_povline"

df_replacements_rows<- data.frame("text" = c("1", "2"),
                                  "slug_suffix" = c("1_00", "2_00"),
                                  "math" = c("1", "2"))



kable(applymyReplaceFunRowwise(df_block = df_block,
                            sheet_name = sheet_name,
                            df_replacements_rows = df_replacements_rows))


```


Now lets run applymyReplaceFunRowwise iteratively. In the first iteration, it runs across the list containing the control stubs. The output of this first iteration is a list of (multirow) dataframes which is the control stubs multiplied up by only one set of Aux values. Further iterations then run across this list of dataframes, which itself outputs a list of dataframes. And so on, for each Aux sheet.

```{r}
for(aux_sheet in aux_sheets){
  
  df_replacements_rows<- read.xlsx(xlsx_url, sheet = aux_sheet)
    
    running_list<- lapply(running_list, applymyReplaceFunRowwise, sheet_name = aux_sheet, df_replacements_rows =  df_replacements_rows)
   
}

# collapse the output into a single df
controls_multiplied_up<- bind_rows(running_list)

# drop any duplicate rows (these will be created where a control stub row doesn't make use of all Aux sheets)
controls_multiplied_up<-  controls_multiplied_up %>% distinct()


```




For each row, for each Aux1, for each Aux2... etc, make a new row swapping in Aux values
```{r}

xlsx_url <- "https://raw.githubusercontent.com/owid/notebooks/main/JoeHasell/Copy_of_Michalis_work/control_stubs.xlsx"

control_stubs_sheet_name<- "control_stubs"


wb<- loadWorkbook(xlsx_url)

# Load the stubs as a dataframe
control_stubs<- read.xlsx(xlsx_url, sheet = control_stubs_sheet_name)

# a vector of the names of the other sheets (the Aux vars)
aux_sheets<- names(wb)
aux_sheets<- aux_sheets[aux_sheets != "control_stubs"]


# Make a list of rows in control_stubs

nested_list<- list()

nested_list[["control_stubs"]]<- split(control_stubs, seq(nrow(control_stubs)))


 #Use lapply with additional functions to move to the next aux block


  
  
  
  
  return(df_block_replaced)
  
}


# Function to go through an aux sheet and multiply up a dataframe with the options in each row of the aux sheet. Argument x will be the dataframe (this will be the element of the list the funciton will be applied to). Argument aux_sheet is a string that is the name of the aux_sheet.
myfun<- function(df_block, aux_sheet){
  
  df_block<- nested_list[["control_stubs"]][[1]]
  aux_sheet<- "abs_povline"
  
  # make a list out of the rows of the auxsheet
  
  df_aux_sheet<- read.xlsx(xlsx_url, sheet = aux_sheet)
  
  list_aux_sheet<- split(df_aux_sheet, seq(nrow(df_aux_sheet)))
  
  # apply across list replacing elements in the dataframe passed to function
  replaced_list<- lapply(list_aux_sheet, myReplaceFun, =df_block)
  
  list of cols
    each row
        replace_with_str<- df_test_aux[aux_list_row, col_name]
  

      search_for_str<- paste0("[", aux_list, "{", col_name, "}]")
  
      # Replace str across whole dataframe
       data.frame(lapply(data, function(x) {gsub(search_for_str, replace_with_str, x)}))
  
    
}

  
for (aux_sheet in aux_sheets) {
nested_list[[aux_sheet]]<- lapply(nested_list[[1]], myfun, aux_sheet = aux_sheet)


  }

# collapse highest order of list into a single df






-----
# An empty dataframe to append the completed set of controls for each row of control_stubs
full_controls<- control_stubs[0,]

# For each row in control_stubs...
for(i in 1:nrow(control_stubs)){

  i<- 1
  
  # Make a list of the Aux vars used in this row
  aux_items_this_row<- as.list(strsplit(control_stubs[i,"Aux"], ',\\s')[[1]])
  
  # Make a dataframe of just the one row of control_stubs in question
  controls_to_repeat<-  control_stubs[i,]
  
  for(aux_list in aux_items_this_row){
    
    controls_to_repeat
  }
  
  aux_list<- "abs_povline"
  for(aux_element in c("1_00", "5_00")){
    
    
    
  }
  
  # grab column names
  
  # for each column col_name ....
  
    # for each row aux_list_row
  
    replace_with_str<- df_test_aux[aux_list_row, col_name]
  

    search_for_str<- paste0("[", aux_list, "{", col_name, "}]")
  
  # Replace str with value
  data <- data.frame(lapply(data, function(x) {gsub(search_for_str, replace_with_str, x)}))
  
  
  # append to running dataframe
  
df_test_aux<- data.frame("text" = c("1", "5"),
                         "slug_suffix"= c("1_00", "5_00"))  
  
  
  pov_lines<- 
  
  test<- expand.grid(controls_to_repeat, pov_lines)
  
  
  }

```



-----

Separate out key/id variables
```{r}
df_key_vars<- df %>%
  select(Entity, Year, isinterpolated, IsSurveyYear, coveragetype, datatype)
```

Separate out absolute poverty line specific vars in long format
```{r}
df_abs_pov<- cbind(df_key_vars, df[, grep("per_day", names(df))]) 

df_abs_pov<- df_abs_pov %>%
  pivot_longer(cols = starts_with("X_"),
               names_to = "old_varname",
               values_to = "value") %>%
  mutate(abs_pov_line_text = gsub("_per_day.*","",old_varname)) %>%
  mutate(abs_pov_line_text = gsub("X_","",abs_pov_line_text))
  mutate(var = gsub(".*_{3}","",old_varname)) %>%
  select(-old_varname)

# Pivot wider
df_abs_pov<- df_abs_pov %>%
  pivot_wider(names_from = var,
              values_from = value)


```

Calculate new variables on the absolute poverty data
```{r}
# Absolute poverty gap per day
df_abs_pov<- df_abs_pov %>%
mutate(absolute_poverty_gap_per_day = absolute_poverty_gap/365)

# Average (absolute) shortfall per day
df_abs_pov<- df_abs_pov %>%
mutate(average_shortfall_per_day = absolute_poverty_gap_per_day/total_number_of_people_below_poverty_line)


```



A set of poverty lines to refer to the poverty-line specific variables
```{r}


abs_pov_lines<- data.frame("rough_pov_line" = unique(df_abs_pov$abs_pov_line_text))
                           
abs_pov_lines<- abs_pov_lines %>%
  mutate(math_pov_line = gsub("_",".",rough_pov_line)) %>%
  mutate(math_pov_line = as.numeric(math_pov_line)) %>%
  mutate(clear_pov_line = paste0("$",str_trim(format(round(math_pov_line, 2), nsmall = 2)), " per day"))
  

 
# 
#                        data.frame(rough_pov_line =  c("1_9",
#                     "3_2",
#                     "5_5",
#                     "10",
#                     "15",
#                     "20",
#                     "30",
#                     "40"),
#                           clean_pov_line = c( "$1.90 a day",
#                     "$3.20 a day",
#                     "$5.50 a day",
#                     "$10 a day",
#                     "$15 a day",
#                     "$20 a day",
#                     "$30 a day",
#                     "$40 a day"))

```




## Controls and data for poverty explorer
```{r}
poverty_metrics<- data.frame(metric_dropdown = c("Share in poverty",
                            "Number of people in poverty",
                            "Average shortfall to poverty line",
                            "Income gap ratio",
                            "Poverty gap index",
                            "Total shortfall to poverty line"),
                     metric_slug_snippet = c("share_of_population_below_poverty_line",
                                               "total_number_of_people_below_poverty_line",
                                             "average_shortfall_per_day",
                                             "income_gap_ratio",
                                             "poverty_gap_index",
                                             "absolute_poverty_gap_per_day"))
```

Absolute poverty controls

```{r}




poverty_metrics<- poverty_metrics %>%
  mutate(title_snippet = c("Poverty: Share of population living on less than",
                           "Poverty: Number of people living on less than",
                            "Depth of poverty: Average daily shortfall below",
                           "Depth of poverty: Income gap ratio at",
                           "Poverty gap index at",
                            "Total daily shortfall below"))


abs_pov_controls<- expand.grid("metric_dropdown" = poverty_metrics$metric_dropdown,
                               "rough_pov_line" = unique(df_abs_pov$abs_pov_line_text))


clear_pov_line<-  paste0("$", format(round(abs_pov_controls$rough_pov_line, 2), nsmall = 2), " per day")

abs_pov_controls<- abs_pov_controls %>%
  mutate(clear_pov_line = paste0("$", format(round(rough_pov_line, 2), nsmall = 2), " per day"))

abs_pov_controls <- left_join(abs_pov_controls, poverty_metrics) %>%
  mutate(ySlugs = paste0(rough_pov_line, "_per_day_", metric_slug_snippet),
         title = paste0(title_snippet, " $", ) 


```

Relative poverty headcount controls
```{r}


pov_lines<- data.frame(rough_pov_line =  c("X_40__of_median_income",
                    "X_50__of_median_income",
                    "X_60__of_median_income"),
                          clean_pov_line = c("40% of median",
                    "50% of median",
                    "60% of median"))


poverty_metrics<- poverty_metrics %>%
  mutate(title_snippet = c("Relative poverty: Share of population living on less than",
                           "Relative poverty: Number of people living on less than",
                           "Depth of relative poverty: Average daily shortfall to",
                           "Depth of relative poverty: Average annual shortfall to",
                           "Total daily shortfall to a poverty line of",
                           "Total annual shortfall to a poverty line of",
                           "Depth of relative poverty: Income gap ratio at",
                           "Poverty gap index at"))



rel_pov_controls<- expand.grid("metric_dropdown" = poverty_metrics$metric_dropdown,
                               "rough_pov_line" = pov_lines$rough_pov_line) 

rel_pov_controls <- left_join(rel_pov_controls, pov_lines) %>%
  mutate(pov_line_dropdown = paste0("Relative line: ", clean_pov_line))

rel_pov_controls <- left_join(rel_pov_controls, poverty_metrics) %>%
  mutate(ySlugs = paste0(rough_pov_line, "___", metric_slug_snippet),
         title = paste0(title_snippet, " ", clean_pov_line))


```

Join controls for the different headcount controls
```{r}
poverty_controls<- bind_rows(abs_pov_controls, rel_pov_controls) %>%
  select(title, metric_dropdown, pov_line_dropdown, ySlugs)

```


Function to repeat controls for each survey data type
```{r}


repeat_controls_by_survey_type<- function(controls_df){

controls_by_survey_type<- list()

controls_by_survey_type[["controls_inc_or_cons"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Either income or consumption",
         extrapolate_checkbox = "'true",
         tableSlug = "inc_or_cons")


controls_by_survey_type[["controls_inc_or_cons_no_extrap"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Either income or consumption",
         extrapolate_checkbox = "'false",
         tableSlug = "inc_or_cons_no_extrap")


controls_by_survey_type[["controls_inc_only"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Income only",
         extrapolate_checkbox = "'true",
         tableSlug = "inc")

controls_by_survey_type[["controls_inc_only_no_extrap"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Income only",
         extrapolate_checkbox = "'false",
         tableSlug = "inc_no_extrap")



controls_by_survey_type[["controls_cons_only"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Consumption only",
         extrapolate_checkbox = "'true",
         tableSlug = "cons")


controls_by_survey_type[["controls_cons_only_no_extrap"]]<- controls_df %>%
  mutate(survey_type_dropdown = "Consumption only",
         extrapolate_checkbox = "'false",
         tableSlug = "cons_no_extrap")




return(controls_by_survey_type)

}

```

Repeat controls for poverty
```{r}
poverty_controls_by_survey_type<- repeat_controls_by_survey_type(poverty_controls)
```


Join survey type controls together
```{r}
poverty_controls_all_survey_types<- bind_rows(poverty_controls_by_survey_type)

poverty_controls_all_survey_types<- poverty_controls_all_survey_types %>%
  select(title, 
         metric_dropdown, 
         pov_line_dropdown, 
         survey_type_dropdown, 
         extrapolate_checkbox, 
         tableSlug, 
         ySlugs)

poverty_controls_all_survey_types<- poverty_controls_all_survey_types %>%
  mutate(yAxisMin = 0)

names(poverty_controls_all_survey_types)<- c("title",
                              "Metric Dropdown", 
                              "Poverty line Dropdown", 
                              "Household survey data type Radio", 
                              "Show extrapolated estimates for years with no survey data Checkbox", 
                              "tableSlug", 
                              "ySlugs",
                              "yAxisMin")


```




Write poverty explorer controls to csv
```{r}
write.csv(poverty_controls_all_survey_types, "poverty_explorer_controls.csv", row.names = F)
```



## Incomes across the distribution explorer controls

```{r}

# thresholds<- c("P10 (poorest tenth)",
#               "P20",
#               "P30",
#               "P40",
#               "Median – P50",
#               "P60",
#               "P70",
#               "P80",
#               "P90 (richest tenth)",
#               clean<- "1","2","3","4","5",
#                         "6", "7", "8", "9")
# 
# deciles<- c("Decile 1 (poorest tenth)",
#             "Decile 2",
#             "Decile 3",
#             "Decile 4",
#             "Decile 5",
#             "Decile 6",
#             "Decile 7",
#             "Decile 8",
#             "Decile 9",
#             "Decile 10 (richest tenth)")
# 

```


## Controls for Inequality explorer
```{r}

```





##Split dfs by survey type



Add annotation for survey type (where this changes over time, say 'Income data before X, consumption after')
```{r}


```




```{r}

# Income only, including extrapolated years
inc<- df %>%
  filter(datatype == "income")

write.csv(inc, "data/inc.csv", row.names = FALSE)

# Income only, excluding extrapolated years
inc_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE,
         datatype == "income")

write.csv(inc_no_extrap, "data/inc_no_extrap.csv", row.names = FALSE)

  
# Consumption only, including extrapolated years
cons<- df %>%
  filter(datatype == "consumption")

write.csv(cons, "data/cons.csv", row.names = FALSE)

# Consumption only, excluding extrapolated years
cons_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE,
         datatype == "consumption")  

write.csv(cons_no_extrap, "data/cons_no_extrap.csv", row.names = FALSE)


# Both income and consumption, including extrapolated years
write.csv(df, "data/inc_or_cons.csv", row.names = FALSE)


# Both income and consumption, excluding extrapolated years
inc_or_cons_no_extrap<- df %>%
  filter(IsSurveyYear == TRUE)

# Check for years where there are both income and consumption surveys

write.csv(inc_or_cons_no_extrap, "data/inc_or_cons_no_extrap.csv", row.names = FALSE)
  




```



