---
title: "OMM Public Health Expenditure"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(jsonlite)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidyr)
library(zoo)

```

Downloading the data from OECD Stat - the data is available [here](https://data.oecd.org/healthres/health-spending.htm).

```{r download}
year <- format(Sys.Date(),"%Y")
oecd_stat_fp <- paste0("data/oecd_stat_",year,".csv")
if (!file.exists(oecd_stat_fp)){
  download.file('https://stats.oecd.org/sdmx-json/data/DP_LIVE/.HEALTHEXP.../OECD?contentType=csv&detail=code&separator=comma&csv-lang=en', oecd_stat_fp )
}

```

Loading in the country standardisation csv
```{r}
countries <- read.csv("data/country_names_country_standardized.csv") %>% distinct()

```



Reading in the data downloaded from OECD Stat. We are using the values relating to 'Government/Compulsory' expenditure - which is here coded 'COMPULSORY' in the 'SUBJECT' column.

In this code chunk we also standardise the country names to OWID country names.

We also add rows so that for each country every year between the first and last has its own row. 
```{r}
oecd <- read.csv(oecd_stat_fp) %>% filter(SUBJECT == 'COMPULSORY' & MEASURE == 'PC_GDP') %>%
  select(entity = LOCATION, year = TIME,value = Value ) %>% 
  mutate(source = 'oecd_stat')%>% 
  left_join(., countries, by = c('entity'= 'Country')) %>% 
  select(entity = Our.World.In.Data.Name, year, value, source)%>% 
  group_by(entity) %>% 
  complete(year = full_seq(min(year):max(year), 1))

ggplot()+
  geom_line(data =oecd, aes(x = year, y = value, group = interaction(entity)))+
  labs(x = "Year", y= "Public Health Expenditure (%GDP)")+
  facet_wrap(.~entity)+
  scale_x_continuous(breaks=seq(1970, 2020, 25))+
  theme_bw()

```


Reading in the data from OECD 1993, which is available [here](https://drive.google.com/file/d/1Xa9WRtQq_Dfeod8mfJzcj6TGXak7APRF/view?usp=sharing.). 

In this code chunk we also standardise the country names to OWID country names and filter out countries that aren't in the OECD stat dataset.

```{r}
oecd_93 <- read.csv("data/OECD 1993.csv") %>% 
  pivot_longer(starts_with('X')) %>% 
  mutate(entity = year, year = as.numeric(gsub("X","",name)), source = 'oecd_93') %>% 
  left_join(., countries, by = c('entity'= 'Country')) %>% 
  select(entity = Our.World.In.Data.Name, year, value, source)%>% 
  filter(entity %in% oecd$entity)  %>% 
  group_by(entity) %>% 
  complete(year = full_seq(min(year):max(year), 1))

ggplot()+
  geom_line(data =oecd_93, aes(x = year, y = value, group = interaction(entity)))+
  labs(x = "Year", y= "Public Health Expenditure (%GDP)")+
  facet_wrap(.~entity)+
  scale_x_continuous(breaks=seq(1960, 1990, 15))+
  theme_bw()
```


Reading in the data from Lindert (1994) - The Rise of Social Spending 1880-1930 (Table 1D), available [here](https://drive.google.com/file/d/1YFlIC-on7oWiDhgdETUD9PSVE-BTaySn/view?usp=sharing).

In this code chunk we also standardise the country names to OWID country names and filter out countries that aren't OECD members.

```{r}
lindert <- read.csv("data/lindert_1880_1930.csv") %>% 
  left_join(., countries, by = c('entity'= 'Country')) %>% 
  mutate(source = 'Lindert') %>% 
  select(entity = Our.World.In.Data.Name, year, value = public_expenditure_on_health, source)%>% 
  select(entity,year,value, source )

ggplot()+
  geom_line(data =lindert, aes(x = year, y = value, group = interaction(entity)))+
  labs(x = "Year", y= "Public Health Expenditure (%GDP)")+
  facet_wrap(.~entity)+
  scale_x_continuous(breaks=seq(1880, 1930, 25))+
  theme_bw()

```

Here we linearly interpolate missing data from both OECD datasets. For example the data for France between is only available every five years between 1970 and 1990, so we interpolate the missing values. 

```{r}
oecd <- oecd %>% group_by(entity) %>%
  mutate(value = na.approx(value, na.rm = FALSE)) %>%
  ungroup()

oecd_93 <- oecd_93 %>% group_by(entity) %>%
  mutate(value = na.approx(value, na.rm = FALSE)) %>%
  ungroup()

```

Here we combine both the datasets and plot them, coloured according to the source. Values which have been linearly interpolated (as described above) are shown as 'OWID'.
```{r}


oecd_both<- rbind(oecd, oecd_93) 
oecd_both$source[is.na(oecd_both$source)] <- 'OWID'
ggplot()+
  geom_line(data =oecd_both, aes(x = year, y = value, group = interaction(source,entity), colour = source), size = 1)+
  scale_x_continuous(breaks=seq(1960, 2020, 30))+
  facet_wrap(.~ entity)

```

Calculating the rate of change in the OECD Stat data. Here we create a new column (time_m1) which is lagged by one year, for example values in the 'value' column for 1971 rows are shown as 1972 in the 'time_m1' column. 

We then the divide the 'time_m1' column by the 'value' column to estimate the rate of change - in a backwards direction. So we would estimate the rate of change from 1972 to 1971.
```{r}
roc_df<- oecd %>% 
  select(entity, year, value) %>% 
  group_by(entity) %>% 
  arrange(entity,year) %>% 
  mutate(time_m1 = lag(value, n = 1L), br_roc = time_m1/value, source = 'oecd_stat') %>% 
  ungroup() %>% 
  filter(complete.cases(.))

head(roc_df)
```

Calculating the rate of change in the OECD 1993 data. Here we create a new column (time_m1) which is lagged by one year, for example values in the 'value' column for 1971 rows are shown as 1972 in the 'time_m1' column. 

We then the divide the 'time_m1' column by the 'value' column to estimate the rate of change - in a backwards direction. So we would estimate the rate of change from 1972 to 1971.
```{r}
oecd_93_roc <- oecd_93 %>% 
  select(entity, year, value) %>% 
  group_by(entity) %>% 
  mutate(time_m1 = lag(value, n = 1L), br_roc = time_m1/value, source = 'oecd_93') %>% 
  ungroup() %>% 
  filter(complete.cases(.))


```



We combine the two rates of change - we can check the rate of change has been calculated as expected by looking at the values in the table below. 

The 1962 value for Australia is 2.74, we multiply this by the 1962 value of the rate of change (br_roc), which is 0.9379562 to get the 1961 value of 2.57.
```{r}
roc_both <- rbind(oecd_93_roc, roc_df)
head(roc_both)
```

As there is some overlap in the time-series there are duplicate for some entity-rate_of_change pairs. In these cases we have a preference for OECD Stat.

```{r}
roc_both <- roc_both %>% 
  left_join(., countries, by = c('entity'= 'Country')) %>% 
  select(entity = Our.World.In.Data.Name, year, br_roc, source)

roc_both <- roc_both %>% 
  group_by(entity, year) %>% 
  add_count() %>%
  ungroup() 

roc_sel <- roc_both[-which(roc_both$n >1 & roc_both$source == 'oecd_93'),]

head(roc_sel)

```

There are some countries which have a gap between the OECD datasets (Belgium, Switzerland). For these we linearly interpolate the missing rate of change values. 

```{r}

roc_int <- roc_sel %>% 
  group_by(entity) %>% 
  complete(year = full_seq(min(year):max(year), 1))%>% 
  mutate(br_roc = na.approx(br_roc, na.rm = FALSE)) %>% 
  ungroup()

head(roc_int)
```
These interpolated values are labelled as 'OWID'

```{r}
roc_int$source[is.na(roc_int$source)] <- 'OWID'

```




This loop iterates through each entity and within this iterates through each year (going backwards through time), to multiply the current 'value' by the current 'br_roc' (rate of change) to calculate the 'value' for the next year.


The calculations for each entity are combined as the loop runs. 
```{r}


oecd_out <- oecd %>%
  select(entity, year, value)
oecd_loop <- NULL
entities <- unique(oecd_out$entity)
for (entity_sel in entities) {
  oecd_sel <- oecd_out %>%
    filter(entity == entity_sel)
  start_reg_year <- min(oecd_sel$year)
  min_roc_year <-
    roc_int %>% group_by(entity) %>%  filter(entity == entity_sel, year == min(year)) %>% ungroup() %>% select(year)  %>% pull()
  
  if ( length(min_roc_year)  == 1) {
    if (min_roc_year <= start_reg_year){
    years <- start_reg_year:min_roc_year
    for (year_sel in years) {
      roc_yr <- roc_int %>% filter(year == year_sel & entity == entity_sel)
      
      oecd_new <- oecd_sel %>% filter(year == year_sel) %>%
        mutate(value = value * roc_yr$br_roc, year = year_sel - 1) %>%
        select(entity, year, value)
      
      oecd_sel <- rbind(oecd_sel, oecd_new)
      
    }
    }
  }
  oecd_loop <- rbind(oecd_loop, oecd_sel)
}

```

Lastly, we add in the Lindert data.
```{r}
full_owid <- rbind(oecd_loop, lindert %>% select(-source))
full_owid <- full_owid %>% select(entity, year, public_health_expenditure_pc_gdp = value) %>% filter(complete.cases(.))

full_owid$source <- 'OWID'
```

Plot the combined dataset to check it looks sensible:

```{r}

ggplot()+
  geom_line(data = full_owid, aes(x = year, y = public_health_expenditure_pc_gdp, group = entity, colour = source), size = 0.8)+
  geom_line(data =oecd_both, aes(x = year, y = value, group = interaction(source,entity), colour = source), size = 1.2, alpha = 0.5)+
  facet_wrap(.~entity)+
  scale_x_continuous(breaks=seq(1880, 2020, 70))+
  theme_bw()
  
```

Write out the data.
```{r}
full_owid <- full_owid %>% select(-source)
write.csv(full_owid,paste0('data/OMM_public_health_expenditure_',year,'.csv'), row.names = FALSE)
```
