---
title: "Cleaning WHO WASH data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(dplyr)

```

## Reading in the data


```{r}
df <- read.csv("data/WASH_data_to_be_cleaned.csv")
```

Identifying which columns should be rounded to integers - columns with 'number' in their name
```{r}

num_cols <- colnames(df)[grepl("number",colnames(df))]
oth_cols <- colnames(df)[!grepl("number",colnames(df))]
oth_cols <- oth_cols[3:length(oth_cols)]

```

Rounding columns to appropriate decimal places. Lots of the negative values were very small and could be solved by rounding. 
```{r}
df <- df %>% 
  mutate_at(num_cols, as.numeric) %>% 
  mutate_at(num_cols, round, 0) %>% 
  mutate_at(oth_cols, as.numeric) %>% 
  mutate_at(oth_cols, round, 2)
```
A handful of columns still had some negative values
```{r}
df$san_imp_number_without[df$san_imp_number_without < 0 & !is.na(df$san_imp_number_without)] <- 0
df$wat_imp_number_without[df$wat_imp_number_without < 0 & !is.na(df$wat_imp_number_without)] <- 0
df$wat_sm_number_without[df$wat_sm_number_without < 0 & !is.na(df$wat_sm_number_without)] <- 0
df$wat_bas_number_without[df$wat_bas_number_without < 0 & !is.na(df$wat_bas_number_without)] <- 0
df$san_bas_number_without[df$san_bas_number_without < 0 & !is.na(df$san_bas_number_without)] <- 0

```

Replacing NAs with empty space for grapher. 
```{r}
df[,3:ncol(df)][is.na(df[,3:ncol(df)])] <- ""

```

Writing out data for upload to grapher. 
```{r}
write.csv(df, "data/Water_and_sanitation_WHO_wash_2021_clean.csv", row.names = FALSE)

```
